---
# The book's home page. Edit the title/subtitle/description to introduce
# yourself and what this portfolio contains.
title: "Welcome to my Data Science Workspace"
author: "Aude GUERAND"
date: today
---

```{r}
#| include: false
# Package setup lives in one shared chunk so every chapter starts from a
# consistent environment. Add libraries you use across the book here.
library(tidyverse)
```

## About Me & This Portfolio 👋

Welcome to my personal data science portfolio! I am **Aude GUERAND**, currently a student in the **Master for Smart Data Science (MSD)** at **ENSAI** (Ecole Nationale de la Statistique et de l'Analyse de l'Information). 

This space is built as part of the *Reproducible Workflows* course. Over the academic year, it will grow into a comprehensive collection of my data science projects, highlighting rigorous engineering standards and fully reproducible analytical pipelines.

Every project hosted here follows modern industry best practices:
- **Reproducible Sourcing**: Automated data pipelines pulling from open-data portals (e.g., data.gouv.fr) directly via code.
- **Environment Isolation**: Strict dependency management pinned using cutting-edge tools like `uv`/`uvr` and `{renv}`.
- **Version Control**: Full project transparency and history trackability hosted on GitHub.

---

## Portfolio Chapters & Projects

This workspace is structured to track my academic progression and engineering outputs:

- **Chapter 1 — Exploratory Data Analysis (EDA)**: My upcoming primary project focused on autonomous data fetching, cleaning, advanced visualization, and insight generation from a public dataset.
- **Chapters 2 & 3 — Core Milestones**: Dynamic placeholders designed to incorporate advanced modeling and data pipelines as the Master's program progresses.

```{r}
#| label: fig-eda-sneak-peek
#| fig-cap: "Initial data glimpse (Temporary placeholder visualization ahead of the EDA project)."
# A tiny placeholder plot so the book renders before you add your own analysis.
ggplot(data.frame(x = rnorm(200), y = rnorm(200)), aes(x, y)) +
  geom_point(alpha = 0.5, colour = "#1A5F7A") + # Synchronized with your custom cobalt blue brand color
  labs(x = "Variable X", y = "Variable Y")
```

---

## Language-Agnostic Infrastructure

Following the core philosophy of a modern data stack, this pipeline is **language-agnostic by design**. While the initial engineering relies on R and the Tidyverse, the underlying deployment architecture—orchestrated with Quarto, versioned through Git, and published to Posit Connect—seamlessly scales across R, Python, or hybrid environments.

::: {.callout-note}
## Smart Data Engineering
As a Smart Data Science student, my focus bridges advanced statistical modeling with software engineering practices to deploy stable, reproducible, and production-ready data systems.
:::
