---
title: "Chapter 1 · EDA · Live coral cover across New Caledonia's reef sites"
subtitle: "As part of the Reproducible Workflows course"
author: "Aude GUERAND"
date: today
---

> **How to fill this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> findings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: find a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: finish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The question

<!-- TODO: In a sentence or two, what question are you trying to answer with this dataset?  -->
What does live hard coral cover look like across the reef sites monitored by
the **RORC** (Réseau d'Observation des Récifs Coralliens), New Caledonia's
volunteer coral reef monitoring network, and how do the sites compare to one
another?

## Getting the data


The data is sourced **by code** from New Caledonia's open-data portal,
[data.gouv.nc](https://data.gouv.nc), published by the RORC network (Agence
calédonienne de la biodiversité and provincial partners). Two related
datasets are used: reef **habitat/substrate** observations and
**invertebrate** counts, both collected on the same stations and transects.

- Habitats: <https://data.gouv.nc/api/v2/catalog/datasets/rorc_habitats_recifaux/exports/csv>
- Invertebrates: <https://data.gouv.nc/api/v2/catalog/datasets/rorc_invertebres/exports/csv>

Both are fetched directly from their stable CSV export endpoints (the
portal's own API), so re-rendering this document always pulls the same data
by construction — no manual download step is involved.

**A note on how the monitoring works.** Each dataset is organised around
four nested levels: a **`site`** (a broad reef location, e.g. "Koumac") is
visited during a **`campagne`** (a monitoring campaign/survey round); within
a site, divers survey one or more **`station`**s (a fixed monitoring spot,
returned to on every campaign so results are comparable over time); and at
each station, observations are recorded along one or more **`transect`**s —
a fixed-length line (commonly 20 or 50 m) laid along the reef, under which
the diver records what lies directly beneath it at regular intervals (coral,
sand, algae, etc.) or counts organisms within a fixed band alongside it.
The transect is the actual unit of observation: it is what turns "what does
this reef look like" into a repeatable, comparable measurement.

```{r}
#| label: setup
#| message: false
library(readr)      # read_csv2() for the semicolon-separated exports
library(janitor)    # clean_names(), tabulate with tabyl()
library(dplyr)      # data wrangling
library(naniar)     # visualise missing values
library(skimr)      # one-line data summary with skim()
library(gt)         # summary table
library(ggplot2)    # the main figure
```
```{r}
#| label: download
#| message: false
url_habitats    <- "https://data.gouv.nc/api/v2/catalog/datasets/rorc_habitats_recifaux/exports/csv"
url_invertebres <- "https://data.gouv.nc/api/v2/catalog/datasets/rorc_invertebres/exports/csv"

habitats_raw    <- read_csv2(url_habitats)
invertebres_raw <- read_csv2(url_invertebres)

habitats    <- habitats_raw    |> clean_names()
invertebres <- invertebres_raw |> clean_names()
```

> **Optionally pin the fetched copy** with `{pins}` so the chapter reads from a
> versioned local board instead of hitting the network on every render:
>
> ```r
> board <- pins::board_folder("pins/", versioned = TRUE)
> pins::pin_write(board, habitats, "rorc_habitats", type = "csv")
> habitats <- pins::pin_read(board, "rorc_habitats")
> ```


## A first look

```{r}
#| label: skim
#| fig-cap: "A one-line summary of the habitats dataset."
skim(habitats)
```

```{r}
#| label: missingness
#| fig-cap: "Missing values across variables."
# vis_miss(dat) — where is data missing, and why might that be?
```

## Cleaning decisions {#sec-cleaning}

<!-- TODO: document the decisions you made while cleaning the data. What did you
rename, recode, or drop, and why? A defensible cleaning decision is a feature
of a good analysis, not an afterthought. -->



Column names were standardised with `janitor::clean_names()` on import.

The `type_substrat` column uses a Reef Check / GCRMN-style benthic
nomenclature rather than plain-language labels (confirmed by inspecting
`distinct(habitats, type_substrat)`):

| Code  | Meaning                    | Kept for "live coral"? |
|-------|-----------------------------|:-----------------------:|
| `HCB` | Hard Coral Branching         | ✅ |
| `HCM` | Hard Coral Massive            | ✅ |
| `HCO` | Hard Coral Other                | ✅ |
| `HCT` | Hard Coral Tabulate               | ✅ |
| `SC`  | Soft Coral                         | ❌ |
| `DC`  | Dead Coral                          | ❌ |
| `FS`  | Fleshy Seaweed (algae)                | — |
| `SP`  | Sponge                                  | — |
| `RC`  | Rock                                      | — |
| `RB`  | Rubble                                     | — |
| `SD`  | Sand                                        | — |
| `SI`  | Silt                                         | — |
| `OT`  | Other                                        | — |

Only the four hard-coral codes (`HCB`, `HCM`, `HCO`, `HCT`) were kept to
isolate **live hard coral**. Two related categories were deliberately
**excluded**: `DC` (Dead Coral, since it is not "live" by definition) and
`SC` (Soft Coral, since soft coral is not part of the reef-building hard
coral community this question focuses on). This is a judgment call — a
version of this analysis including soft coral would answer a slightly
different question, and is a natural extension noted in Findings.


```{r}
#| label: clean-filter
coral_labels <- c("HCB", "HCM", "HCO", "HCT")

coral_habitats <- habitats |>
  filter(type_substrat %in% coral_labels)

coral_cover_by_site <- coral_habitats |>
  group_by(site) |>
  summarise(
    total_coral_count = sum(decompte, na.rm = TRUE),
    n_observations = n(),
    .groups = "drop"
  ) |>
  arrange(desc(total_coral_count))
```

## Plotting the story

<!-- TODO: replace with your figures. Give every figure a caption and refer to
it in the text (e.g. "Fig. 1 shows that …"). If a figure uses randomness
(sampling, jitter, a simulation), fix the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the figure is identical on every render. -->


```{r}
#| label: tbl-coral-cover
#| tbl-cap: "Live hard coral cover by site."
coral_cover_by_site |>
  gt() |>
  tab_header(
    title = "Live coral cover by site",
    subtitle = "RORC New Caledonia — Réseau d'Observation des Récifs Coralliens"
  ) |>
  cols_label(
    site = "Site",
    total_coral_count = "Total coral count",
    n_observations = "Number of observations"
  )
```

```{r}
#| label: fig-main
#| fig-cap: "Total live hard coral count by site, RORC New Caledonia."
coral_cover_by_site |>
  ggplot(aes(x = reorder(site, total_coral_count), y = total_coral_count)) +
  geom_col(fill = "#2C7A7B") +
  coord_flip() +
  labs(
    x = "Site",
    y = "Total live hard coral count",
    title = "Live hard coral cover by site"
  ) +
  theme_minimal()
```

Fig. 1 shows Koumac with the highest total coral count, followed by
Poindimie, Deva, La Foa, Ouegoa and Port Bouquet, in decreasing order.


## Findings

<!-- TODO: what did you learn? This is your discussion: put your findings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT find. -->


Koumac has both the highest total coral count (1916) and the most
observations recorded (544), while Port Bouquet has the fewest of both (687
and 192 respectively). This raises a methodological caveat: the raw totals
in @tbl-coral-cover partly reflect **sampling effort**, not just reef
condition. A fairer, effort-adjusted comparison would look at
`total_coral_count / n_observations` per site rather than the raw sum — a
natural next step for this chapter.

What this analysis does **not** establish: it is a single-snapshot summary
across the years present in the export, so no trend over time is shown here,
and no explanation is offered for *why* some sites carry more live coral
than others (exposure, depth, distance to river mouths, past bleaching
events, etc. are all plausible but untested here). Extending the analysis to
use the `campagne` field to look at coral cover over time, or joining with
the `invertebres` table (on `campagne` + `site` + `station` + `transect`) to
see whether coral-rich sites also host greater invertebrate diversity, are
both natural directions for a later revision of this chapter.



---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the final gallery walk.*
